Efficiently stable presentations from error-correcting codes

### 2024-02-26 arXiv_v2/efficient_stability.tex
added arXiv_v2 with small bug fix (data-processing)

\begin{abstract}
We introduce a notion of \emph{efficient stability} for finite presentations of groups. Informally, a finite presentation using generators $S$ and relations $R$ is \emph{stable} if any map from $S$ to unitaries 
%(say, in some finite von Neumann algebra) 
that approximately satisfies the relations (in the tracial norm) is close to the restriction of a representation of $G$ to the subset $S$. This notion and variants thereof have been extensively studied in recent years, in part motivated by connections to property testing in computer science. The novelty in our work is the focus on \emph{efficiency}, which, informally, places an onus on small presentations --- in the sense of encoding length.  
%(of size, say, logarithmic in the group size)
The goal in this setup is to achieve non-trivial tradeoffs between the presentation length and its modulus of stability.
%---as, intuitively, more generators and relations place more constraints on the approximate homomorphism, making stability easier to achieve---whereas we wish to minimize the number of variables and constraints.

@@

We will make use of the following simple fact.

\begin{lemma}[Data-processing]\label{lem:dp}
Let $\{P_a\}$ and $\{Q_a\}$ be two POVMs on $(\mM,\tau)$ with the same outcome set $\mA$. Then for any function $f:\mA\to \mB$ for some finite set $\mB$, 
\begin{equation}\label{eq:dp}
\sum_{b\in \mB} \Big\| \sum_{a\in f^{-1}(b)} (P_a-Q_a) \Big\|_2^2 \,\leq\, \sum_{a\in \mA} \big\| P_a-Q_a\big\|_2^2\;.
\end{equation}
\end{lemma}

\begin{proof}
This follows by expanding the left-hand side and using $\tau(P_aQ_{a'})\geq 0$ for all $a\neq a'$. 
\end{proof}

@@

Applying Lemma~\ref{lem:pull-back}, we obtain projective measurements $\{Q^W_u\}$ on $\mM$ such that 
\begin{equation}\label{eq:main-1}
\sum_u \big\|Q^W_u - (w^W)^* \tilde{P}_u^W w^W \big\|^2_2 \,=\, O(\delta_1)\;.
\end{equation}
For any $i\in\{1,\ldots,n\}$ let $Q^{W,i}_b = \sum_{u:\,u_i=b}  Q^W_u$. Then by Lemma~\ref{lem:dp},
\begin{align*}
\Es{i}\sum_b \tau\big( Q^{W,i}_b (w^W)^*\tilde{P}^{W,i}_b(w^W) \big)
&\geq \Es{i}\sum_u \tau\big( Q^{W}_u (w^W)^*\tilde{P}^{W}_u(w^W) \big)\\
&\geq 1-O(\delta_1)\;.
\end{align*}
For $W\in\{X,Z\}$ and $b\in \F_2^k$ let $R^W_b = Q^W_{G_\mC^T b}$, where by definition $G_\mC^T b\in \mC$.

>>>>

\begin{abstract}
We introduce a notion of \emph{efficient stability} for finite presentations of groups. Informally, a finite presentation using generators $S$ and relations $R$ is \emph{stable} if any map from $S$ to unitaries 
that approximately satisfies the relations (in the tracial norm) is close to the restriction of a representation of $G$ to the subset $S$. This notion and variants thereof have been extensively studied in recent years, in part motivated by connections to property testing in computer science. The novelty in our work is the focus on \emph{efficiency}, which, informally, places an onus on small presentations --- in the sense of encoding length.  
%(of size, say, logarithmic in the group size)
The goal in this setup is to achieve non-trivial tradeoffs between the presentation length and its modulus of stability.
%---as, intuitively, more generators and relations place more constraints on the approximate homomorphism, making stability easier to achieve---whereas we wish to minimize the number of variables and constraints.

@@

\tnote{removing erroneous dp fact}
%We will make use of the following simple fact.

%\begin{lemma}[Data-processing]\label{lem:dp}
%Let $\{P_a\}$ and $\{Q_a\}$ be two POVMs on $(\mM,\tau)$ with the same outcome set $\mA$. Then for any function $f:\mA\to \mB$ for some finite set $\mB$, 
%\begin{equation}\label{eq:dp}
%\sum_{b\in \mB} \Big\| \sum_{a\in f^{-1}(b)} (P_a-Q_a) \Big\|_2^2 \,\leq\, \sum_{a\in \mA} \big\| P_a-Q_a\big\|_2^2\;.
%\end{equation}
%\end{lemma}

%\begin{proof}
%This follows by expanding the left-hand side and using $\tau(P_aQ_{a'})\geq 0$ for all $a\neq a'$. 
%\end{proof}

@@

Applying Lemma~\ref{lem:pull-back}, we obtain projective measurements $\{Q^W_u\}$ on $\mM$ such that 
\begin{equation}\label{eq:main-1}
\sum_u \big\|Q^W_u - (w^W)^* \tilde{P}_u^W w^W \big\|^2_2 \,=\, O(\delta_1)\;.
\end{equation}
For any $i\in\{1,\ldots,n\}$ let $Q^{W,i}_b = \sum_{u:\,u_i=b}  Q^W_u$. \tnote{Added comment here to replace erroneous DP fact:}Expanding both terms on the left-hand side and using that $\tau(Q^W_u \tilde{P}^W_{u'})\geq 0$ for all $u,u'$,
\begin{align*}
\Es{i}\sum_b \tau\big( Q^{W,i}_b (w^W)^*\tilde{P}^{W,i}_b(w^W) \big)
&\geq \Es{i}\sum_u \tau\big( Q^{W}_u (w^W)^*\tilde{P}^{W}_u(w^W) \big)\\
&\geq 1-O(\delta_1)\;.
\end{align*}
For $W\in\{X,Z\}$ and $b\in \F_2^k$ let $R^W_b = Q^W_{G_\mC^T b}$, where by definition $G_\mC^T b\in \mC$.
